---
title: "code_prscx_score for multi-ancestry VTE PRS"
author: "Yon Ho Jee"
date: '2022 6 10 '
---
#Steps
#(1)Preparation of the weight file to grab the RSID that identical to the bim file in the genotype file  
#(2)compute PRS using plink 
#(3)sum the PRS score from different chromosome
#(4)Evaluate model PRS
#   Find "meta-weights" for combined PRS (**only applicable to Michigan biobank**)

# (1)Preparation of the weight file to grab the RSID that identical to the bim file in the genotype file 
please save weight files to /path/output/

# (2)compute PRS using plink 

**EUR**
```{bash, eval = FALSE}
module load plink/1.90-fasrc01

for chr in {1..22};do
plink --bfile /n/holylfs05/LABS/kraft_lab/Lab/yjee/VTE/Processed_Data/VTEhumancore_imputed_Rsq095_rsid --score /n/holylfs05/LABS/kraft_lab/Lab/yjee/VTE/Processed_Data/output/weight_EUR_pst_eff_a1_b0.5_phiauto_chr"$chr".txt 2 4 6 sum --out /n/holylfs05/LABS/kraft_lab/Lab/yjee/VTE/Processed_Data/score/score_EUR_pst_eff_a1_b0.5_phiauto_chr"$chr".txt
done
```

**AFR**
```{bash, eval = FALSE}
module load plink/1.90-fasrc01

for chr in {1..22};do
plink --bfile /n/holylfs05/LABS/kraft_lab/Lab/yjee/VTE/Processed_Data/VTEhumancore_imputed_Rsq095_rsid --score /n/holylfs05/LABS/kraft_lab/Lab/yjee/VTE/Processed_Data/output/weight_AFR_pst_eff_a1_b0.5_phiauto_chr"$chr".txt 2 4 6 sum --out /n/holylfs05/LABS/kraft_lab/Lab/yjee/VTE/Processed_Data/score/score_AFR_pst_eff_a1_b0.5_phiauto_chr"$chr".txt
done
```

#(3)Sum the PRS score from different chromosome
```{r}
for(weight in c("AFR","EUR")){
phi = "phiauto"

a = 1 # starting chromosome number
b = 22 #ending chromosome number

output.file.name =paste("/path/score/", "finalscore_", weight, "_pst_eff_a1_b0.5_", phi, "_chr", a,"-chr", b, ".csv", sep ="")

chr = a
score.file.name = paste("/path/score/", "score_", weight, "_pst_eff_a1_b0.5_", phi, "_chr", chr,".txt.profile", sep ="")
score = read.table(score.file.name, header = T)
final.score = score[, c(1:2)]
final.score$PRS = 0

final.score$PRS = final.score$PRS + score$SCORESUM

for (chr in c((a+1):b)){
  
  score.file.name = paste("/path/score/", "score_", weight, "_pst_eff_a1_b0.5_", phi, "_chr", chr,".txt.profile", sep ="")
  score = read.table(score.file.name, header = T)
  final.score$PRS = final.score$PRS + score$SCORESUM
  print(chr)
}

head(final.score)

write.csv(final.score, file = output.file.name)

}

```

```{r}
finalscore_AFR<-read.csv("/path/score/finalscore_AFR_pst_eff_a1_b0.5_phiauto_chr1-chr22.csv")
finalscore_EUR<-read.csv("/path/score/finalscore_EUR_pst_eff_a1_b0.5_phiauto_chr1-chr22.csv")
finalscore <- inner_join(finalscore_AFR, finalscore_EUR, by = c("FID","IID"))
names(finalscore)[4]<-"PRSCSX_AFR";names(finalscore)[6]<-"PRSCSX_EUR"
  
hist(finalscore_AFR$PRS);hist(finalscore_EUR$PRS)

```

# (4)Evaluate model PRS

#load phenotype data ("IID","sex", "PC1-10", "caco" as outcome)
```{r}
y2<-"pheno"
#make sure that phenotype data are in the same order as bed file
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(qqman)
library(pROC)
library(tidyverse)
```

# Standardize to controls
```{r}
y3 <- inner_join(y2, finalscore, by = c("IID")) #4423

y3$PRSCSX_AFR_std<-(y3$PRSCSX_AFR-mean(y3$PRSCSX_AFR[y3$caco==0]))/sd(y3$PRSCSX_AFR[y3$caco==0])
y3$PRSCSX_EUR_std<-(y3$PRSCSX_EUR-mean(y3$PRSCSX_EUR[y3$caco==0]))/sd(y3$PRSCSX_EUR[y3$caco==0])
  
hist(y3$PRSCSX_AFR_std);hist(y3$PRSCSX_EUR_std)

```

### Find "meta-weights" for combined PRS  - **only applicable to Michigan biobank**
```{r}
weight_reg.formula <- paste("") %>%
    paste0("caco ~ PRSCSX_AFR_std + PRSCSX_EUR_std", .) %>%
    as.formula

train.dat<-y3

prscsx_weight.model <- glm(weight_reg.formula, data = train.dat, family = "binomial") %>% 
  summary
prscsx_weight.model
mylogit <- glm(weight_reg.formula, data = train.dat, family = "binomial") 

coef(mylogit)

# (Intercept) PRSCSX_AFR_std PRSCSX_EUR_std 
#    <b0>         <b1>        <b3> 
# please provide b0, b1, b3 and plug these betas for LPRS_combined_AFR calculation below
```

PRSCSX_combined_EUR
```{r}
y3$PRSCSX_combined_EUR <- 0.06614*y3$PRSCSX_AFR_std  + 0.34631*y3$PRSCSX_EUR_std -0.07715 
y3$PRSCSX_combined_EUR_std<-(y3$PRSCSX_combined_EUR-mean(y3$PRSCSX_combined_EUR[y3$caco==0]))/sd(y3$PRSCSX_combined_EUR[y3$caco==0])

prscx_combined_eur.formula <- paste("PC", 1:10, sep = "", collapse = "+") %>%
    paste0("caco ~ PRSCSX_combined_EUR_std+sex+", .) %>%
    as.formula

test.dat<-y3

mylogit_combined_eur <- glm(prscx_combined_eur.formula, data = test.dat, family = "binomial") 
```

PRSCSX_combined_AFR
```{r}
y3$PRSCSX_combined_AFR <- 0.1193056*y3$PRSCSX_AFR_std  + 0.1776201*y3$PRSCSX_EUR_std -2.7424880
y3$PRSCSX_combined_AFR_std<-(y3$PRSCSX_combined_AFR-mean(y3$PRSCSX_combined_AFR[y3$caco==0]))/sd(y3$PRSCSX_combined_AFR[y3$caco==0])

prscx_combined_afr.formula <- paste("PC", 1:10, sep = "", collapse = "+") %>%
    paste0("caco ~ PRSCSX_combined_AFR_std+sex+", .) %>%
    as.formula

test.dat<-y3

mylogit_combined_afr <- glm(prscx_combined_afr.formula, data = test.dat, family = "binomial") 

```

PRSCSX_EUR
```{r}
prscx_eur_reg.formula <- paste("PC", 1:10, sep = "", collapse = "+") %>%
    paste0("caco ~ PRSCSX_EUR_std + sex +", .) %>%
    as.formula

mylogit_eur <- glm(prscx_eur_reg.formula, data = test.dat, family = "binomial") 
```

PRSCSX_AFR
```{r}
prscx_afr_reg.formula <- paste("PC", 1:10, sep = "", collapse = "+") %>%
    paste0("caco ~ PRSCSX_AFR_std + sex +", .) %>%
    as.formula
mylogit_afr <- glm(prscx_afr_reg.formula, data = test.dat, family = "binomial") 
```

### Make ROC curves, compute AUC

```{r}
or1<-exp(cbind(OR = coef(mylogit_eur), confint(mylogit_eur)))[2,]
or2-exp(cbind(OR = coef(mylogit_afr), confint(mylogit_afr)))[2,]
or3<-exp(cbind(OR = coef(mylogit_combined_eur), confint(mylogit_combined_eur)))[2,]
or4<-exp(cbind(OR = coef(mylogit_combined_afr), confint(mylogit_combined_afr)))[2,]
```

```{r}
or_std_cx <-  as.data.frame(matrix(rep(NA,16),ncol = 4, nrow = 4))
colnames(or_std_cx) <- c("Method","OR","OR_Low","OR_High")
or_std_cx[1,1] <- "Stdandarized PRSCSx-EUR"
or_std_cx[2,1] <- "Stdandarized PRSCSx-AFR"
or_std_cx[3,1] <- "Stdandarized PRSCSx-combined_EUR"
or_std_cx[4,1] <- "Stdandarized PRSCSx-combined_AFR"

or_std_cx$OR[1] <- or1[1]
or_std_cx$OR_Low[1] <- or1[2]
or_std_cx$OR_High[1] <- or1[3]

or_std_cx$OR[2] <- or2[1]
or_std_cx$OR_Low[2] <- or2[2]
or_std_cx$OR_High[2] <- or2[3]

or_std_cx$OR[3] <- or3[1]
or_std_cx$OR_Low[3] <- or3[2]
or_std_cx$OR_High[3] <- or3[3]

or_std_cx$OR[4] <- or4[1]
or_std_cx$OR_Low[4] <- or4[2]
or_std_cx$OR_High[4] <- or4[3]

kable(or_std_cx[1:4,])
kbl(or_std_cx[1:4,] , caption = "PRSCSx") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

```{r}
results <-  as.data.frame(matrix(rep(NA,16),ncol = 4, nrow = 4))
colnames(results) <- c("Method","AUC","AUC_Low","AUC_High")
results[1,1] <- "PRSCSx EUR"
results[2,1] <- "PRSCSx AFR"
results[3,1] <- "PRSCSx combined_EUR"
results[4,1] <- "PRSCSx combined_AFR"

test_roc_eur = roc(test.dat$caco ~ test.dat$PRSCSX_EUR_std)
results$AUC[1] <- auc(test_roc_eur)
results$AUC_Low[1] <- ci.auc(test_roc_eur)[1]
results$AUC_High[1] <- ci.auc(test_roc_eur)[3]

test_roc_afr = roc(test.dat$caco ~ test.dat$PRSCSX_AFR_std)
results$AUC[2] <- auc(test_roc_afr)
results$AUC_Low[2] <- ci.auc(test_roc_afr)[1]
results$AUC_High[2] <- ci.auc(test_roc_afr)[3]

test_roc_combined_EUR = roc(test.dat$caco ~ test.dat$PRSCSX_combined_EUR_std)
results$AUC[3] <- auc(test_roc_combined_EUR)
results$AUC_Low[3] <- ci.auc(test_roc_combined_EUR)[1]
results$AUC_High[3] <- ci.auc(test_roc_combined_EUR)[3]

test_roc_combined_AFR = roc(test.dat$caco ~ test.dat$PRSCSX_combined_AFR_std)
results$AUC[4] <- auc(test_roc_combined_AFR)
results$AUC_Low[4] <- ci.auc(test_roc_combined_AFR)[1]
results$AUC_High[4] <- ci.auc(test_roc_combined_AFR)[3]

kable(results[1:4,])
kbl(results[1:4,] , caption = "PRSCSx") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

