---
title: "Code LDpred2_auto for multi-ancestry VTE PRS"
author: "Yon Ho Jee"
date: '2022 6 10 '
---

#Steps
#(1)load bed file & SNP list & beta_auto
#(2)Apply PRS to VTE individual-level data
#(3)Evaluate model PRS
#   Find "meta-weights" for combined PRS (**only applicable to Michigan biobank**)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(qqman)
library(pROC)
library(bigstatsr)
library(parallel)
library(kableExtra)
library(snpStats)
library(caret)
library(gridExtra)
library(MetaSubtract)
library(tidyverse)
library(R.utils)
```

```{r}
library(bigsnpr)
options(bigstatsr.check.parallel.blas = FALSE)
options(default.nproc.blas = NULL)
```

# (1)load bed file & SNP list & beta_auto

### Get VTE bed/bim/fam files as bigSNP object
```{r}
NCORES <- parallelly::availableCores()
tmpVTE <- tempfile(tmpdir = "scratch")
on.exit(file.remove(paste0(tmpVTE, ".sbk")), add = TRUE)

path2 <- snp_readBed2("VTEhumancore_imputed_Rsq095.bed", backingfile = tmpVTE)

obj.bigSNPVTE <- snp_attach(path2)

str(obj.bigSNPVTE)

dim(obj.bigSNPVTE$genotypes)

```

### Imputation
```{r}
genotypeVTE <- snp_fastImputeSimple(
obj.bigSNPVTE$genotypes,
method = "mode",
ncores = NCORES
)
```

# load SNP list
```{r}
info_snp_new_EUR<-readRDS("info_snp_EUR")
info_snp_new_AFR<-readRDS("info_snp_AFR")
```

```{r}
# extract the SNP information from the genotype
mapVTE <- obj.bigSNPVTE$map[-3]
names(mapVTE) <- c("chr", "rsid", "pos", "a1", "a0")

# perform SNP matching
info_snpVTE_EUR <- snp_match(info_snp_new_EUR, mapVTE)
df_beta_EUR <- info_snpVTE_EUR[, c("pos","_NUM_ID_")]

info_snpVTE_AFR <- snp_match(info_snp_new_AFR, mapVTE)
df_beta_AFR <- info_snpVTE_AFR[, c("pos","_NUM_ID_")]
```

# load beta_auto
```{r}
beta_auto_EUR<-readRDS("beta_auto_EUR.Jun_10_2022") #604741     48
beta_auto_AFR<-readRDS("beta_auto_AFR.Apr_03_2022") #1184805      48
```

# (2)Apply PRS to VTE individual-level data

LDpred2_EUR
```{r}
# calculate PRS for all samples
ind.test <- 1:nrow(genotypeVTE) 
pred_auto <- 
    big_prodMat(genotypeVTE,
                beta_auto_EUR,
                ind.row = ind.test, 
                ind.col = df_beta_EUR$`_NUM_ID_`) 
# scale the PRS generated from AUTO
pred_scaled <- apply(pred_auto, 2, sd) 
final_beta_auto_EUR <- #sum across multiple models (columns) 
    rowMeans(beta_auto_EUR[,
                abs(pred_scaled -
                    median(pred_scaled)) <
                    3 * mad(pred_scaled)])
pred_auto <-
    big_prodVec(genotypeVTE,
        final_beta_auto_EUR,
        ind.row = ind.test,
        ind.col = df_beta_EUR$`_NUM_ID_`)

sum(final_beta_auto_EUR !=0) #604741
```

LDpred2_AFR
```{r}
# calculate PRS for all samples
ind.test <- 1:nrow(genotypeVTE) #4423
pred_auto_AFR <- #4423 4
    big_prodMat(genotypeVTE,
                beta_auto_AFR[df_beta_AFR$`_NUM_ID_`,],
                ind.row = ind.test, #4423
                ind.col = df_beta_AFR$`_NUM_ID_`) #191172
# scale the PRS generated from AUTO
pred_scaled_AFR <- apply(pred_auto_AFR, 2, sd) #4
final_beta_auto_AFR <- #sum across multiple models (columns) -> #191172
    rowMeans(beta_auto_AFR[,
                abs(pred_scaled_AFR -
                    median(pred_scaled_AFR)) <
                    3 * mad(pred_scaled_AFR)])
pred_auto_AFR <-
    big_prodVec(genotypeVTE,
        final_beta_auto_AFR[df_beta_AFR$`_NUM_ID_`],
        ind.row = ind.test,
        ind.col = df_beta_AFR$`_NUM_ID_`)

sum(final_beta_auto_AFR !=0) #1184805
```

# (3)Evaluate model PRS

#load phenotype data ("sex", "PC1-10", "caco" as outcome)
```{r}
y2<-"pheno"
#make sure that phenotype data are in the same order as bed file
```

#Standardize PRS
```{r}
reg.dat <- y2
reg.dat$LPRS_EUR <- pred_auto 
reg.dat$LPRS_AFR <- pred_auto_AFR 

reg.dat$LPRS_EUR_std <- (reg.dat$LPRS_EUR-mean(reg.dat$LPRS_EUR[reg.dat$caco==0]))/sd(reg.dat$LPRS_EUR[reg.dat$caco==0])
reg.dat$LPRS_AFR_std <- (reg.dat$LPRS_AFR-mean(reg.dat$LPRS_AFR[reg.dat$caco==0]))/sd(reg.dat$LPRS_AFR[reg.dat$caco==0])

reg_EUR_std.formula <- paste("PC", 1:10, sep = "", collapse = "+") %>%
    paste0("caco ~ LPRS_EUR_std+sex+", .) %>%
    as.formula
reg_AFR_std.formula <- paste("PC", 1:10, sep = "", collapse = "+") %>%
    paste0("caco ~ LPRS_AFR_std+sex+", .) %>%
    as.formula
```

### Find "meta-weights" for combined PRS  - **only applicable to Michigan biobank**
```{r}
weight_reg.formula <- paste("") %>%
    paste0("caco ~ LPRS_AFR_std + LPRS_EUR_std", .) %>%
    as.formula

train.dat<-reg.dat

ldpred2_weight.model <- glm(weight_reg.formula, data = train.dat, family = "binomial") %>% 
  summary
ldpred2_weight.model
mylogit <- glm(weight_reg.formula, data = train.dat, family = "binomial") 

coef(mylogit)

# (Intercept) LPRS_AFR_std LPRS_EUR_std 
#    <b0>         <b1>        <b3> 
# please provide b0, b1, b3 and plug these betas for LPRS_combined_AFR calculation below
```

LDpred2_combined_EUR
```{r}
reg.dat$LPRS_combined_EUR <- 0.02819114*reg.dat$LPRS_AFR_std + 0.38007277*reg.dat$LPRS_EUR_std -0.07654566

reg.dat$LPRS_combined_EUR_std<-(reg.dat$LPRS_combined_EUR-mean(reg.dat$LPRS_combined_EUR[reg.dat$caco==0]))/sd(reg.dat$LPRS_combined_EUR[reg.dat$caco==0])

ldpred2_combined_EUR_reg.formula <- paste("PC", 1:10, sep = "", collapse = "+") %>%
    paste0("caco ~ LPRS_combined_EUR_std+sex+", .) %>%
    as.formula
```

LDpred2_combined_AFR
```{r}
reg.dat$LPRS_combined_AFR <- 0.09164556*reg.dat$LPRS_AFR_std + 0.21466434*reg.dat$LPRS_EUR_std -2.74059333

reg.dat$LPRS_combined_AFR_std<-(reg.dat$LPRS_combined_AFR-mean(reg.dat$LPRS_combined_AFR[reg.dat$caco==0]))/sd(reg.dat$LPRS_combined_AFR[reg.dat$caco==0])

ldpred2_combined_AFR_reg.formula <- paste("PC", 1:10, sep = "", collapse = "+") %>%
    paste0("caco ~ LPRS_combined_AFR_std+sex+", .) %>%
    as.formula
```

### Make ROC curves, compute AUC

```{r}
model1 <- glm(reg_EUR_std.formula, data = reg.dat, family = "binomial") 
model2 <- glm(reg_AFR_std.formula, data = reg.dat, family = "binomial") 
model3 <- glm(ldpred2_combined_EUR_reg.formula, data = reg.dat, family = "binomial") 
model4 <- glm(ldpred2_combined_AFR_reg.formula, data = reg.dat, family = "binomial") 

or1<-exp(cbind(OR = coef(model1), confint(model1)))[2,]
or2<-exp(cbind(OR = coef(model2), confint(model2)))[2,]
or3<-exp(cbind(OR = coef(model3), confint(model3)))[2,]
or4<-exp(cbind(OR = coef(model4), confint(model4)))[2,]
```

```{r}
or_std_lp <-  as.data.frame(matrix(rep(NA,12),ncol = 4, nrow = 3))
colnames(or_std_lp) <- c("Method","OR","OR_Low","OR_High")
or_std_lp[1,1] <- "Stdandarized LDpred2-EUR"
or_std_lp[2,1] <- "Stdandarized LDpred2-AFR"
or_std_lp[3,1] <- "Stdandarized LDpred2-combined_EUR"
or_std_lp[4,1] <- "Stdandarized LDpred2-combined_AFR"

or_std_lp$OR[1] <- or1[1]
or_std_lp$OR_Low[1] <- or1[2]
or_std_lp$OR_High[1] <- or1[3]

or_std_lp$OR[2] <- or2[1]
or_std_lp$OR_Low[2] <- or2[2]
or_std_lp$OR_High[2] <- or2[3]

or_std_lp$OR[3] <- or3[1]
or_std_lp$OR_Low[3] <- or3[2]
or_std_lp$OR_High[3] <- or3[3]

or_std_lp$OR[4] <- or4[1]
or_std_lp$OR_Low[4] <- or4[2]
or_std_lp$OR_High[4] <- or4[3]

kable(or_std_lp[1:4,])
kbl(or_std_lp[1:4,] , caption = "LDpred2") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

```{r}
results <-  as.data.frame(matrix(rep(NA,16),ncol = 4, nrow = 4))
colnames(results) <- c("Method","AUC","AUC_Low","AUC_High")
results[1,1] <- "LDpred2 EUR"
results[2,1] <- "LDpred2 AFR"
results[3,1] <- "LDpred2 combined_EUR"
results[4,1] <- "LDpred2 combined_AFR"

test_roc_eur = roc(reg.dat$caco ~ reg.dat$LPRS_EUR_std)
results$AUC[1] <- auc(test_roc_eur)
results$AUC_Low[1] <- ci.auc(test_roc_eur)[1]
results$AUC_High[1] <- ci.auc(test_roc_eur)[3]

test_roc_afr = roc(reg.dat$caco ~ reg.dat$LPRS_AFR_std)
results$AUC[2] <- auc(test_roc_afr)
results$AUC_Low[2] <- ci.auc(test_roc_afr)[1]
results$AUC_High[2] <- ci.auc(test_roc_afr)[3]

test_roc_combined_EUR = roc(reg.dat$caco ~ reg.dat$LPRS_combined_EUR)
results$AUC[3] <- auc(test_roc_combined_EUR)
results$AUC_Low[3] <- ci.auc(test_roc_combined_EUR)[1]
results$AUC_High[3] <- ci.auc(test_roc_combined_EUR)[3]

test_roc_combined_AFR = roc(reg.dat$caco ~ reg.dat$LPRS_combined_AFR)
results$AUC[4] <- auc(test_roc_combined_AFR)
results$AUC_Low[4] <- ci.auc(test_roc_combined_AFR)[1]
results$AUC_High[4] <- ci.auc(test_roc_combined_AFR)[3]

kable(results[1:4,])
kbl(results[1:4,] , caption = "LDpred2") %>%
  kable_classic(full_width = F, html_font = "Cambria")

```

